---
title: "Data_Analysis"
author: "Sara Milla - 12432969"
date: "2026-06-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Setup
```{r}
setwd("C:/Users/saram/OneDrive/COURSES/DASM")
library(tidyverse)
library(stringr)
library(ggplot2)
library(dplyr)

if (!requireNamespace("janitor", quietly = TRUE)) {
  install.packages("janitor")
}

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("survey", quietly = TRUE)) {
  install.packages("survey")
}
if (!requireNamespace("car", quietly = TRUE)) {
  install.packages("car")
}

library(janitor)
library(patchwork)
library(car)

load_orig_dataset <- function(){
  
  all_results = read.csv("Survey_results_2026.csv")
  dim(all_results)
  colnames(all_results)
  
  return (all_results)
}

create_g13_subset <- function(all_results){
  
  df = all_results %>% select(starts_with("DQ") | starts_with("G13"))
  colnames(df) = c("age", "gender", "academic_program", "Q1", "Q2", "Q3")
  
  return(df)
}


show_rundown <- function(df){
  dim(df)
  str(df)
  head(df)
}

all_results <- load_orig_dataset()
df <- create_g13_subset(all_results)
show_rundown(df)
```

General Exploration before Question-specific Analysis

```{r}
df %>% 
  tabyl(academic_program, gender) %>% 
  adorn_totals(where = c("row", "col"))

gender_over_age_hist <- df %>% 
  count(age, gender) %>% 
  ggplot(aes(x = age, y = n, fill = gender)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Gender", 
       x = "Age Group", y = "Count", fill = "Gender") + 
  theme_minimal()

prog_over_age_hist <- df %>% 
  count(age, academic_program) %>% 
  ggplot(aes(x = age, y = n, fill = academic_program)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Academic Program", 
       x = "Age Group", y = "Count", fill = "Program") + 
  theme_minimal()

# Displaying distribution of gender and program over age groups
gender_over_age_hist + prog_over_age_hist
```

We observe that Data Science students span a wider range of age categories compared to Business Informatics, where no students from the age groups 18-20 or over 35 participated in the survey. Data Science has a larger share of students with 177 compared to 32 for Business Informatics.Furthermore, the distribution of female students has a different shapecompared to the overall distribution, as the size of the 21-23 age groupexceeds that of the 23-25 group for females. As male students, accounting for 156 out of 209 observations of the sample, appear more in 23-25 compared to 21-23, they dominate the overall shape too.

#Question 1
After 10 years of successful safety testing, on a scale of 1 to 5 (1 = Strongly Disagree, 5 = Strongly Agree), would you undergo a surgical procedure to implant a cognitive-enhancement chip in your brain?

```{r}
table(df$Q1)
round(prop.table(table(df$Q1)), 4)
barplot(prop.table(table(df$Q1)), ylim = c(0,0.4))
```
#Question 2
After 10 years of successful safety testing, what ratio of the general population (0-100%) do you believe would choose to get the implant? 
```{r}
table(df$Q2)
round(prop.table(table(df$Q2))*100, 2)
hist(df$Q2, ylim = c(0, 60))
hist(df$Q2,  freq = FALSE, ylim = c(0, 0.03))
```

# Q2 analysis
```{r}
# Boxplot of Q2 (Estimated Ratio) by Academic Program
ggplot(df, aes(x = academic_program, y = Q2, fill = academic_program)) +
  geom_boxplot() +
  labs(title = "Estimated Implant Adoption by Academic Program",
       x = "Academic Program",
       y = "Estimated Population Ratio (0-100%)") +
  theme_minimal()

# Create a summary table grouped by Academic Program
q2_summary <- df %>%
  group_by(academic_program) %>%
  summarise(
    Count = n(),
    Mean = round(mean(Q2, na.rm = TRUE), 2),
    Median = median(Q2, na.rm = TRUE),
    SD = round(sd(Q2, na.rm = TRUE), 2),
    Min = min(Q2, na.rm = TRUE),
    Max = max(Q2, na.rm = TRUE)
  )

print(q2_summary)

# Fit a survey-weighted linear model for Q2
q2_weighted_model <- svyglm(Q2 ~ academic_program, design = final_design)

# Get the rigorous summary (coefficients, standard errors, and p-values)
summary(q2_weighted_model)

```
Looking at gender

```{r}
# Boxplot of Q2 by Academic Program AND Gender
ggplot(df, aes(x = academic_program, y = Q2, fill = gender)) +
  geom_boxplot() +
  labs(title = "Estimated Implant Adoption by Program and Gender",
       x = "Academic Program",
       y = "Estimated Population Ratio (0-100%)",
       fill = "Gender") +
  theme_minimal()

# Fit a survey-weighted linear model for Q2 with an interaction term
q2_interaction_model <- svyglm(Q2 ~ academic_program * gender, design = final_design)

# View the results
summary(q2_interaction_model)
```



